---
title: "Preliminary Models"
author: "Sebastian Rojas Cabal"
date: "7/7/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prelim, message = FALSE, echo= FALSE}
library(tidyverse)
library(stargazer)
library(gridExtra)
library(lmtest)
library(sandwich)
library(lfe)
library(plm)
```

```{r loading_data, echo=FALSE, message = FALSE}
nonprofits_analysis <- read_csv("/Users/srojascabal/Desktop/000_f990_data/analytical_sample_220705.csv",
     col_types = cols(
      ein = col_character(),
      tax_year = col_character(),
      anti_lgbtq = col_double(),
      anti_factor = col_factor(),
      rtrn_state = col_factor(),
      totalXpns_2013_100k = col_double(),
      frgnXpns_2013_100k = col_double(),
      log_frgnXpns_2013_100k = col_double(),
      propFrgnXpns_2013_100k = col_double(),
      yearMrgEq_rtrn = col_double(),
      ind_yearMrgEq_rtrn = col_double(),
      gdp_state_2012 = col_double(),       
      gov_party = col_factor(),
      gov_republican = col_double(),
      F990_990EZ = col_double(),        
      state_religiosity = col_double())
      ) %>%
  mutate(
    log_gdp_2012 = log(gdp_state_2012),
    log_gdp_2012_100k = log(gdp_state_2012_100k),
    log_frgnXpns_2013_100k = log(frgnXpns_2013_100k+1),
    log_totalXpns_2013_100k = log(totalXpns_2013_100k+1),
    log_state_religiosity = log(state_religiosity + 1),
    log_F990_990EZ = log(F990_990EZ + 1)
  )
```

```{r descriptives, results = "asis", message = FALSE, echo= FALSE}
stargazer(as.data.frame(nonprofits_analysis), type = 'latex')
```

```{r plots_distribution_logVars, echo=FALSE}
# GDP
plot_gdp_2012 <-
  ggplot(nonprofits_analysis, aes(x = gdp_state_2012)) +
  geom_density() +
  labs(
    title = "GDP by State"
  )

plot_log_gdp_2012 <-
  ggplot(nonprofits_analysis, aes(x = log_gdp_2012)) +
  geom_density() +
  labs(
    title = "Logged GDP by State"
  )

# GDP 100k
plot_gdp_2012_100k <-
  ggplot(nonprofits_analysis, aes(x = gdp_state_2012_100k)) +
  geom_density()  +
  labs(
    title = "GDP by State (100k USD)"
  )

plot_log_gdp_2012_100k <-
  ggplot(nonprofits_analysis, aes(x = log_gdp_2012_100k)) +
  geom_density() +
  labs(
    title = "Logged GDP by State (100k USD)"
  )

# Foreign Expenses
plot_frgnXpns_2013_100k <-
  ggplot(nonprofits_analysis, aes(x = frgnXpns_2013_100k)) +
  geom_density()  +
  labs(
    title = "Foreign Expenses"
  )

plot_log_frgnXpns_2013_100k <-
  ggplot(nonprofits_analysis, aes(x = log_frgnXpns_2013_100k)) +
  geom_density() +
  labs(
    title = "Logged Foreign Expenses"
  )

# Total Expenses
plot_totalXpns_2013_100k <-
  ggplot(nonprofits_analysis, aes(x = totalXpns_2013_100k)) +
  geom_density() +
  labs(
    title = "Total Expenses"
  )

plot_log_totalXpns_2013_100k <-
  ggplot(nonprofits_analysis, aes(x = log_totalXpns_2013_100k)) +
  geom_density() +
  labs(
    title = "Logged Total Expenses"
  )

# State Religiosity
plot_state_religiosity <-
  ggplot(nonprofits_analysis, aes(x = state_religiosity)) +
  geom_density() +
  labs(
    title = "Number of Churches and Religious Orgs"
  )

plot_log_state_religiosity <-
  ggplot(nonprofits_analysis, aes(x = log_state_religiosity)) +
  geom_density() +
  labs(
    title = "Logged State Religiosity"
  )

# 990/EZ filers
plot_F990_990EZ <-
  ggplot(nonprofits_analysis, aes(x = F990_990EZ)) +
  geom_density() +
  labs(
    title = "Number of F990/EZ Filing Orgs"
  )

plot_log_F990_990EZ <-
  ggplot(nonprofits_analysis, aes(x = log_F990_990EZ)) +
  geom_density() +
  labs(
    title = "Logged Nonprofit Activity"
  )
```

```{r figures_gdp, echo=FALSE}
grid.arrange(plot_gdp_2012, plot_log_gdp_2012,
             ncol = 2) 
```

```{r figures_gdp_100k, echo=FALSE}
grid.arrange(plot_gdp_2012_100k, plot_log_gdp_2012_100k,
             ncol = 2) 
```

```{r figures_frgnXpns_2013_100k, echo=FALSE}
grid.arrange(plot_frgnXpns_2013_100k, plot_log_frgnXpns_2013_100k,
             ncol = 2) 
```

```{r figures_totalXpns_2013_100k, echo=FALSE}
grid.arrange(plot_totalXpns_2013_100k, plot_log_totalXpns_2013_100k,
             ncol = 2) 
```

```{r figures_state_religiosity, echo=FALSE}
grid.arrange(plot_state_religiosity, plot_log_state_religiosity,
             ncol = 2) 
```

```{r figures_F990_990EZ, echo=FALSE}
grid.arrange(plot_F990_990EZ, plot_log_F990_990EZ,
             ncol = 2) 
```

## Models without Fixed Effects Robust Standard Errors
```{r models, echo=FALSE}
# Logged Foreign Expenses, no controls
model1 <- lm(log_frgnXpns_2013_100k ~ anti_lgbtq,
             nonprofits_analysis)

# Foreign Expenses, controls + interaction
model2 <- lm(log_frgnXpns_2013_100k ~ anti_lgbtq + 
               anti_lgbtq*ind_yearMrgEq_rtrn +
               gov_republican + 
               log_gdp_2012 + 
               ind_yearMrgEq_rtrn + 
               log_F990_990EZ + 
               log_state_religiosity + 
               log_totalXpns_2013_100k,
             nonprofits_analysis)

# % of Total Expenditures, no controls
model3 <- lm(propFrgnXpns_2013_100k ~ anti_lgbtq,
             nonprofits_analysis)

# % of Total Expenditures, controls + interaction
model4 <- lm(propFrgnXpns_2013_100k ~ anti_lgbtq +
                anti_lgbtq*ind_yearMrgEq_rtrn +
                gov_republican +
                log_gdp_2012 +
                ind_yearMrgEq_rtrn +
                log_F990_990EZ +
                log_state_religiosity,
              nonprofits_analysis)
```

```{r model_results_TotalExpenses, results = "asis", echo=FALSE}
stargazer(model1, model2, type = 'latex',
          title = "Foreign Expenditures by Anti-LGBTQ Nonprofits, 2013-2020",
          covariate.labels = c("Anti-LGBTQ Nonprofit",
                               "Republican Governor",
                               "State real GDP (Log)",
                               "Marriage Equality Passed",
                               "Nonprofit Activity in State (Log)",
                               "Religiosity of State Population (Log)",
                               "Org's Total Expenses (Log)",
                               "Interaction:Anti/Marriage Eq"),
          dep.var.labels   = c("Log of Total Foreign Expenses"),
          single.row = TRUE)
```

```{r model_results_PctnExpenses, results = "asis", echo=FALSE}
stargazer(model3, model4, type = 'latex',
           title = "Foreign Expenses as a Proportion of Total Expenses  by Anti-LGBTQ Nonprofits, 2013-2020",
           covariate.labels = c("Anti-LGBTQ Nonprofit",
                                "Republican Governor",
                                "State real GDP (Log)",
                                "Marriage Equality Passed",
                                "Nonprofit Activity in State (Log)",
                                "Religiosity of State Population (Log)",
                               "Interaction:Anti/Marriage Eq"),
           dep.var.labels   = c("Foreign Expenses as Proportion of Total Expenses"),
           single.row = TRUE)
```

## Models with Fixed Effects and Robust Standard Errors
```{r models_FE, echo=FALSE}
anti_nonprofits <- nonprofits_analysis %>%
  filter(anti_lgbtq == 1)

# Logged Foreign Expenses, no controls
model1_fe <- felm(log_frgnXpns_2013_100k ~ anti_lgbtq |
                  rtrn_state + ein |
                  0 | 
                  rtrn_state + ein,
                  anti_nonprofits)

# Foreign Expenses, controls
model2_fe <- felm(log_frgnXpns_2013_100k ~ anti_lgbtq + gov_republican + log_gdp_2012 + ind_yearMrgEq_rtrn + log_F990_990EZ + log_state_religiosity + log_totalXpns_2013_100k |
                rtrn_state + ein |
                0 | 
                rtrn_state + ein,
                anti_nonprofits)

# % of Total Expenditures, no controls
model3_fe <- felm(propFrgnXpns_2013_100k ~ anti_lgbtq |
                  rtrn_state + ein |
                  0 | 
                  rtrn_state + ein,
                  anti_nonprofits)

# % of Total Expenditures, controls
model4_fe <- felm(propFrgnXpns_2013_100k ~ anti_lgbtq + gov_republican + log_gdp_2012 + ind_yearMrgEq_rtrn + log_F990_990EZ + log_state_religiosity + log_totalXpns_2013_100k |
                rtrn_state + ein |
                0 | 
                rtrn_state + ein,
                anti_nonprofits)

```

# Total Foreign Expenses (Log), no controls
```{r modelFE_results_TotalExpenses, echo=FALSE}
summary(model1_fe)
```

# Total Foreign Expenses (Log), with  controls
```{r modelFE_results_TotalExpenses2, echo=FALSE}
summary(model2_fe)
```

# % of Total Expenses, no controls
```{r modelFE_results_PctnExpenses, echo=FALSE}
summary(model3_fe)
```

# % of Total Expenses, with  controls
```{r modelFE_results_PctnExpenses2, echo=FALSE}
summary(model4_fe)
```

## Figures

### Foreign Expenses as Proportion of Total Expenses by US Nonprofits, 2013-2020
```{r figures_proportion, echo=FALSE, message=FALSE}
nonprofits_analysis %>%
    mutate(
      factor_txyr = as.factor(tax_year)
    ) %>%
    group_by(factor_txyr, anti_factor) %>%
    summarise(
      expenses = mean(propFrgnXpns_2013_100k)
    ) %>%
    ggplot() +
    geom_line(aes(x = factor_txyr, y = expenses, group = anti_factor, color = anti_factor)) +
    geom_point(aes(x = factor_txyr, y = expenses, color = anti_factor)) +
    labs(y = "Avg. Foreign Expenses as \nProportion of All Expenses",
         caption = "N (Non Anti-LGBTQ+): 1,636,682\nN (Anti-LGBTQ+): 5,608\nAmounts in real 2013 USD\nDashed line indicates the year same-sex marriage was legalized at the federal level") +
    theme_bw() +
    theme(legend.position = "bottom",
          axis.title.x=element_blank(),
          legend.title = element_blank()) +
   geom_vline(xintercept = "2015", linetype = "dashed", color = "gray29")

ggsave("/Users/srojascabal/Desktop/000_f990_data/Plots/prop_national.png",
       width = 210,
       height = 145, #A5 (half of A4)
       units = "mm",
       dpi = "print"
)
```

### Gross Foreign Expenses by US Nonprofits, 2013-2020
```{r figures_gross, echo=FALSE, , message=FALSE}
nonprofits_analysis %>%
    mutate(
      factor_txyr = as.factor(tax_year)
    ) %>%
    group_by(factor_txyr, anti_factor) %>%
    summarise(
      expenses = mean(frgnXpns_2013_100k)
    ) %>%
    ggplot() +
    geom_line(aes(x = factor_txyr, y = expenses, group = anti_factor, color = anti_factor)) +
    geom_point(aes(x = factor_txyr, y = expenses, color = anti_factor)) +
    labs(y = "Average Foreign Expenses \n(Hundreds of Thousands of USD)",
         caption = "N (Non Anti-LGBTQ+): 1,636,682\nN (Anti-LGBTQ+): 5,608\nAmounts in real 2013 USD\nDashed line indicates the year same-sex marriage was legalized at the federal level") +
    theme_bw() +
    theme(legend.position = "bottom",
          axis.title.x=element_blank(),
          legend.title = element_blank()) +
   geom_vline(xintercept = "2015", linetype = "dashed", color = "gray29")

ggsave("/Users/srojascabal/Desktop/000_f990_data/Plots/gross_national.png",
       width = 210,
       height = 145, #A5 (half of A4)
       units = "mm",
       dpi = "print"
)
```

### Anti-LGBTQ+ Nonprofits by State, 2013-2020
```{r figures_antiOrgsState, echo=FALSE}
plot_AntiOrgsState <- nonprofits_analysis %>%
  filter(anti_lgbtq == 1) %>%
  count(rtrn_state) %>%
  arrange(n) %>%
  ggplot(aes(y=n, x=rtrn_state)) +
  geom_bar(position="dodge", stat = "identity") +
  coord_flip() +
  scale_fill_grey(start = 0.8, end = 0.2) +
  theme_classic() +
  labs(
    y = "Number of Anti-LGBTQ+ Nonprofits",
    x= ""
  )
```

```{r echo=FALSE}
plot_AntiOrgsState
```

```{r states_gross_plots, echo = FALSE}
states_for_selection <- nonprofits_analysis %>%
  filter(anti_lgbtq == 1) %>%
  count(rtrn_state) %>%
  pull(rtrn_state)

state_anti_n <- nonprofits_analysis %>%
  filter(anti_lgbtq == 1) %>%
  count(rtrn_state) %>%
  pull(n)

state_nonanti_n <- nonprofits_analysis %>%
  filter(anti_lgbtq == 0) %>%
  count(rtrn_state) %>%
  pull(n)

state_mrg_eq_yr <- nonprofits_analysis %>%
  select(rtrn_state, yearMrgEq_rtrn) %>%
  distinct() %>%
  pull(yearMrgEq_rtrn)

for (i in 1:length(states_for_selection)) {
# Defining Elements for Plot
state <- states_for_selection[i]

state_plot_caption <- paste0(
  "N (Non Anti-LGBTQ+): ",
  state_nonanti_n[i],
  "\nN (Anti-LGBTQ+): ",
  state_anti_n[i],
  "\nAmounts in real 2013 USD",
  "\nDashed line indicates the year same-sex marriage was legalized at the state level"
)

state_plot_y_lab_gross <- c("Average Foreign Expenses \n(Hundreds of Thousands of USD)")

state_plot_filename_gross <- paste0(
  "/Users/srojascabal/Desktop/000_f990_data/Plots/",
       "frgn_gross_",
       states_for_selection[i],
       ".png"
)

state_plot_title <- state

state_plot_subtitle <- paste0(
  "Same-sex marriage legalized in ",
  state_mrg_eq_yr[i]
)

# Plot
nonprofits_analysis %>%
    filter(rtrn_state %in% states_for_selection[i]) %>%
    mutate(
      factor_txyr = as.factor(tax_year)
    ) %>%
    group_by(factor_txyr, anti_factor) %>%
    summarise(
      expenses = mean(frgnXpns_2013_100k)
    ) %>%
    ggplot() +
    geom_line(aes(x = factor_txyr, y = expenses, group = anti_factor, color = anti_factor)) +
    geom_point(aes(x = factor_txyr, y = expenses, color = anti_factor)) +
    labs(
      title = state_plot_title,
      subtitle = state_plot_subtitle,
      y = state_plot_y_lab_gross,
      caption = state_plot_caption
        ) +
    theme_bw() +
    theme(legend.position = "bottom",
          axis.title.x=element_blank(),
          legend.title = element_blank()) +
   geom_vline(xintercept = as.character(state_mrg_eq_yr[i]), linetype = "dashed", color = "gray29")

# Saving plot
ggsave(state_plot_filename_gross,
       width = 210,
       height = 145, #A5 (half of A4)
       units = "mm",
       dpi = "print"
)
}
```

```{r states_proportion_plots, echo = FALSE}
states_for_selection <- nonprofits_analysis %>%
  filter(anti_lgbtq == 1) %>%
  count(rtrn_state) %>%
  pull(rtrn_state)

state_anti_n <- nonprofits_analysis %>%
  filter(anti_lgbtq == 1) %>%
  count(rtrn_state) %>%
  pull(n)

state_nonanti_n <- nonprofits_analysis %>%
  filter(anti_lgbtq == 0) %>%
  count(rtrn_state) %>%
  pull(n)

state_mrg_eq_yr <- nonprofits_analysis %>%
  select(rtrn_state, yearMrgEq_rtrn) %>%
  distinct() %>%
  pull(yearMrgEq_rtrn)


for (i in 1:length(states_for_selection)) {
# Defining Elements for Plot
state <- states_for_selection[i]
  
state_plot_y_lab_prop <- c("Avg. Foreign Expenses as \nProportion of All Expenses")
  
state_plot_caption <- paste0(
  "N (Non Anti-LGBTQ+): ",
  state_nonanti_n[i],
  "\nN (Anti-LGBTQ+): ",
  state_anti_n[i],
  "\nAmounts in real 2013 USD",
  "\nDashed line indicates the year same-sex marriage was legalized at the state level"
)  
  
state_plot_title <- state

state_plot_subtitle <- paste0(
  "Same-sex marriage legalized in ",
  state_mrg_eq_yr[i]
)

state_plot_filename_prop <- paste0(
  "/Users/srojascabal/Desktop/000_f990_data/Plots/",
       "frgn_prop_",
       state,
       ".png"
)

# Plot
nonprofits_analysis %>%
    filter(rtrn_state %in% state) %>%  
    mutate(
      factor_txyr = as.factor(tax_year)
    ) %>%
    group_by(factor_txyr, anti_factor) %>%
    summarise(
      expenses = mean(propFrgnXpns_2013_100k)
    ) %>%
    ggplot() +
    geom_line(aes(x = factor_txyr, y = expenses, group = anti_factor, color = anti_factor)) +
    geom_point(aes(x = factor_txyr, y = expenses, color = anti_factor)) +
    labs(
      title = state_plot_title,
      subtitle = state_plot_subtitle,
      y = state_plot_y_lab_prop,
      caption = state_plot_caption
        ) +
    theme_bw() +
    theme(legend.position = "bottom",
          axis.title.x=element_blank(),
          legend.title = element_blank()) +
   geom_vline(xintercept = as.character(state_mrg_eq_yr[i]), linetype = "dashed", color = "gray29")

# Saving plot
ggsave(state_plot_filename_prop,
       width = 210,
       height = 145, #A5 (half of A4)
       units = "mm",
       dpi = "print",
      )
}

```